### 2020-06-10 chapters/sota.tex
..

The authors define a time-series as \textit{"observations that have been recorded in an orderly fashion and which are correlated in time"}. Regarding the concept of \textit{outlier} the authors state that \textit{"outliers can be thought of as observations that do not follow the expected behavior"}. This definition is similar to the concept of outlier provided by Hawkins in 1980 where he states that an outlier is \textit{“an observation which deviates so much from other observations as to arouse suspicions that it was
generated by a different mechanism"} \cite{Hawkins-Outliers}. Furthermore, Blázquez-García et. al allude to the fact that outliers are often referred to in the literature as \textit{"anomalies, discordant observations, discords, exceptions, aberrations, surprises, peculiarities or contaminants"} which for the purpose of their review and this Thesis all refer to outliers. Hence, outlier detection refers to the process of finding anomalies in data. An anomaly relates to a point in time where the behavior of the system does not match the expected one. Note that under this definition an anomaly does not necessarily imply a problem.

@@

The \textit{input data} axis refers to the type of time-series the method uses as input. According to this criterion, an outlier detection method is said to work with a univariate time-series or a multivariate one. A univariate time-series refers to a time-series consisting of a list of scalars generated by a single time-dependent variable. To quote the authors, a univariate time-series is \textit{"an ordered set of real-valued observations, where each observation is recorded at a specific time"}. On the other hand, a multivariate time-series has more than one time-dependent variable. Most importantly, not only is each variable time-dependant but also may have dependencies on other variables. With that in mind, the authors define a multivariate time-series as \textit{"an ordered set of k-dimensional vectors, each of which is recorded at a specific time"}. Each \textit{k-dimensional} vector consists of \textit{k}  time-dependent variables, possibly correlated.

>>>>

The authors define a time-series as \textit{"observations that have been recorded in an orderly fashion and which are correlated in time"}. Regarding the concept of \textit{outlier} the authors state that \textit{"outliers can be thought of as observations that do not follow the expected behavior"}. This definition is similar to the concept of outlier provided by Hawkins in 1980 where he states that an outlier is \textit{“an observation which deviates so much from other observations as to arouse suspicions that it was
generated by a different mechanism"} \cite{Hawkins-Outliers}. Furthermore, Blázquez-García et. al allude to the fact that outliers are often referred to in the literature as \textit{"anomalies, discordant observations, discords, exceptions, aberrations, surprises, peculiarities or contaminants"} which for the purpose of their review and this Thesis all refer to outliers. Hence, outlier detection refers to the process of finding anomalies in data. An anomaly relates to a point in time where the behavior of the system does not match the expected one. Note that under this definition the presence of outliers or anomalies does not necessarily imply a problem.

@@

The \textit{input data} axis refers to the type of time-series the method uses as input. According to this criterion, an outlier detection method is said to work with a univariate time-series or a multivariate one. A univariate time-series refers to a time-series consisting of a list of values generated by a single time-dependent variable throughout time. To quote the authors, a univariate time-series is \textit{"an ordered set of real-valued observations, where each observation is recorded at a specific time"}. On the other hand, a multivariate time-series has more than one time-dependent variable. Most importantly, not only is each variable time-dependant but also may have dependencies on other variables. With that in mind, the authors define a multivariate time-series as \textit{"an ordered set of k-dimensional vectors, each of which is recorded at a specific time"}. Each \textit{k-dimensional} vector consists of \textit{k}  time-dependent variables, possibly correlated.
